\documentclass{tox}

\begin{document}

\title{A secure, distributed multimedia messenger}

\abstract{\
  \textbf{Abstract}:
Tox is a secure, lightweight, distributed messenger platform designed to 
provide maximum security without penalizing performance.
}

\author{Written by irungentoo} %add your name here when you contribute to it.

\maketitle

\tableofcontents
\clearpage

\section{Introduction}

When one first starts a Tox client, a long term asymmetric keypair is created. 
This keypair identifies the user uniquely. If a Tox user wants to connect to 
another user, they need to obtain that user's long term public key. 

\subsection{Crypto}

Tox uses crypto\_box() from the NaCl crypto library for all the cryptography in 
Tox. Unless otherwise noted, all keys refer to keys generated with 
crypto\_box\_keypair(), all encryption is done with crypto\_box() and all 
decryption with crypto\_box\_open(). For performance purposes the functions to 
precompute the shared secret and encrypt and decrypt messages with it are used 
extensively in Tox; however, this is not relevant to this document.

The function crypto\_box() provides fast public-key authenticated encryption. 
For exactly how it works read the NaCl docs.

\subsection{Tox ID}

The Tox ID contains the long term public key of a Tox user with a nospam value and 
a checksum appended to it. The exact format is:

[NaCl crypto\_box() public key (32 bytes)][nospam (4 bytes)][checksum (2 bytes)]

The Tox ID is 38 bytes and usually distributed in hex format. A typical valid 
Tox ID looks like:

56A1ADE4B65B86BCD51CC73E2CD4E542179F47959FE3E0E21B4B0ACDADE51855D34D34D37CB5

The nospam value is there to enable users to quickly change their Tox IDs 
without changing identities. How it works is explained later in the friend 
request section.

The checksum prevents users from adding mistyped or badly copied Tox IDs. It is 
a simple XOR checksum. Code to compute the checksum from a partial ID of 
ID\_length follows.

\begin{lstlisting}
uint8_t checksum[2] = {0};
uint32_t i;

for (i = 0; i < ID_length; ++i)
    checksum[i % 2] ^= ID[i];

\end{lstlisting}

\subsection{Friends}

To add another Tox user as a friend, one must first obtain their Tox ID. The
Tox ID can be transmitted across insecure channels or posted publicly. One 
must then make sure, however, that the Tox ID belongs to the friend and 
not an impersonator.

Each Tox instance contains a friends list. This is a list of friends 
that the user has added to their client. Tox will try to connect to everyone in 
the friends list. In order to connect to a friend, the friend must have the 
user in their friends list or else the connection will fail. Friends to whom the 
user is connected are shown as online, while friends who are not connected are 
shown as offline.

Users who are not in the friends list of a certain user will not be able to connect
to that user.

There are two ways to add an entry to the friends list. The first is to add the 
Tox ID of a Tox user and the second is to accept a friend request.

\subsection{Friend Requests}

When a Tox user adds a Tox ID, Tox sends a friend request to the Tox user 
corresponding to the added ID. A friend request contains the long term public 
key of the user making the request, the nospam contained in the Tox ID and a 
custom user specified message.

When Tox receives a friend request, it checks whether the nospam is valid (i.e. whether 
it matches the one used in the user's current Tox ID). If it isn't valid, the friend 
request is discarded automatically. If it is valid, the friend request is presented to the 
user along with the message specified in it.

The user can then accept the request and add the other user to their friends list 
or they can refuse it and simply discard the friend request.

\subsection{Text Communications}

Tox users have various ways of communicating with online friends. The first is 
basic text communication. Tox uses UTF-8 as the encoding for all text 
transmitted through it. Using UTF-8 provides extensive language support and support for 
unicode smileys and other characters.

Each Tox user has a nickname and a status that can be changed at will. They are 
transmitted to friends every time a connection to them is established and to 
all currently online friends when the user changes it.

Tox messages can be sent as normal messages or as actions (the equivalent of 
IRC's /me).

\subsection{File Transfers}

Tox supports sending files to online friends. Users can pause and resume files 
as they are being transmitted. If a client becomes disconnected while the transfer is 
in progress and then reconnect afterwards, Tox supports resuming the file transfer.

File transfers in Tox are much faster than other typical messaging software due 
to the fact that it connects directly to other peers most of the time.

\subsection{Audio/Video}

Tox users can make audio and video calls to other Tox users. Call settings can 
be modified during the call which means a user can stop sending video or audio 
while a call is in progress. The video resolution can be changed dynamically 
during a video call. Opus is the codec used for audio and VP8 is the codec used 
for video.

\subsection{Conclusion}

This introduction was meant to give the reader a quick high-level overview of 
Tox: the part that directly interacts with the user. The next section will deal 
with how Tox works at a lower level.

\section{Deeper Into Tox}

The Tox core can be broken down into 2 parts: The peer finding part and the 
messenger part. The peer finding part takes care of finding the peer (friend) 
we 
want to establish a secure connection with and the messenger part takes care of
connecting to and communicating with the peer after the peer finding succeeds.

\subsection{Peer Finding}

The peer finding part has the purpose of finding peers inside the network. It 
is the part that finds the friends in the friend list of the Tox client and 
sends them the required information so that they can establish a secure 
connection with each other.

Peer finding has two parts: the onion and the DHT which work closely together.

The Tox DHT is a UDP-only DHT that Tox clients advertise themselves to with a 
temporary public key called the DHT public key. The DHT public key is 
regenerated every time the Tox client is restarted to prevent the tracking of 
Tox clients inside the network.

Because the DHT only works with temporary keys, there needed to be a way to 
communicate to peers when all that was known was their real long term public 
key. This part is named the onion. The onion is used to anonymously send and 
receive friend requests and packets from friends containing what their current 
DHT public key is along with some TCP relays they are currently connected to.

Once the peer knows the DHT public key of a friend, they query the DHT then 
finds and connects to the friend. In case the friend cannot be found in the DHT 
Tox falls back to using TCP relays to connect to the friend by connecting to 
the relays the friend is connected to and using them to relay data to the 
friend.

The peer finding succeeds when the Tox client has found the best way to send 
data to the friend.

\subsection{Messenger}

Once the peer finding succeeds, a secure connection to the friend is 
established using the Tox protocol. The Tox protocol has perfect forward 
secrecy and supports the padding of messages to obscure the length. It is 
designed to be resistant to replay attacks and to prevent attackers from 
preforming man in the middle attacks on it. All data sent with the protocol 
is encrypted with crypto\_box() and is immune to attacks where the attacker 
could modify the encrypted packet to modify the plain text. Exactly how it 
works is explained later in this document.

Unlike TCP, the Tox protocol does not accept a stream of data but instead 
accepts packets of data. Sending both lossy and lossless packets is supported. 
Lossless packets are lossless and are guaranteed to arrive in the order they 
are sent. Lossy packets on the other hand might not arrive in the order they 
are sent, might not even arrive at all or might even arrive more than once 
because unlike lossless packets, they are lossy and therefore have no replay 
protection.

Every data related packet in Tox except audio and video data packets are 
transmitted as lossless packets. Audio and video data packets are transmitted 
as being lossy because low latency is preferred and small amounts of packet 
loss are not an issue.

The first byte of a packet sent over the Tox protocol determines their type. 
Packet numbers 0 to 15 inclusively are reserved for internal use by the Tox 
protocol. Packet numbers 16 to 191 inclusively and 255 are lossless packets. 
Packet numbers 192 to 254 are lossy packets.

What follows is a list of all currently valid Tox lossless packets that are 
sent to friends over the Tox protocol.

%put this in some other kind of format

\begin{tabular}[pos]{|l|p{9cm}|}

\hline
Packet ID       &      16
\\ \hline
Name            &      Alive
\\ \hline
Size (bytes)    &      1
\\ \hline
Format          &      [byte (16)]
\\ \hline
Description     &      Sent every 5 seconds to each friend. If none are 
                       received from a friend for 15 seconds, the connection 
                       times out.
\\ \hline

\end{tabular}

\begin{tabular}[pos]{|l|p{9cm}|}

\hline
Packet ID       &      17
\\ \hline
Name            &      Share Relays
\\
\hline
Size (bytes)    &      38 to 785
\\ \hline
Format          &      [byte (17)][up to 16 nodes in DHT packed format]
                       % add a link to packet format explanation here
\\ \hline
Description     &      Sent every couple of minutes to every friend. They are 
                       the way TCP relays are shared among friends.
\\ \hline

\end{tabular}

\begin{tabular}[pos]{|l|p{9cm}|}

\hline
Packet ID       &      48
\\ \hline
Name            &      Nickname
\\ \hline
Size (bytes)    &      2 to 129
\\ \hline
Format          &      [byte (48)][NickName            (byte array)]
\\ \hline
Description     &      Used to send the nickname of the Tox user to friends. 
                       The nickname is a non null terminated UTF-8 string of 
                       maximum length 128.
\\ \hline

\end{tabular}

\begin{tabular}[pos]{|l|p{9cm}|}

\hline
Packet ID       &      49
\\ \hline
Name            &      Status Message
\\ \hline
Size (bytes)    &      2 to 1008
\\ \hline
Format          &      [byte (49)][Status Message (byte array)]
\\ \hline
Description     &      Used to send the status message of the Tox user to 
                       friends. The status message is a non null terminated 
                       UTF-8 string of maximum length 1007.
\\ \hline

\end{tabular}

\begin{tabular}[pos]{|l|p{9cm}|}

\hline
Packet ID       &      50
\\ \hline
Name            &      User status
\\ \hline
Size (bytes)    &      2
\\ \hline
Format          &      [byte (50)][byte (User status)]
\\ \hline
Description     &      Used to send the user status of the Tox user to friends 
                       (online (0), away (1), busy (2)). 
\\ \hline
\end{tabular}

\begin{tabular}[pos]{|l|p{9cm}|}

\hline
Packet ID       &      51
\\ \hline
Name            &      Typing
\\ \hline
Size (bytes)    &      2
\\ \hline
Format          &      [byte (51)][byte (typing status)]
\\ \hline
Description     &      Used to send a change in typing status of the Tox user 
                       to friends (not typing (0), typing (non zero)). 
\\ \hline

\end{tabular}

\begin{tabular}[pos]{|l|p{9cm}|}

\hline
Packet ID       &      64
\\ \hline
Name            &      Message
\\ \hline
Size (bytes)    &      2 to 1373
\\ \hline
Format          &      [byte (64)][Message (byte array)]
\\ \hline
Description     &      Used to send text messages to friends. The message is 
                       a non null UTF-8 string with a maximum length of 1372 bytes.
\\ \hline

\end{tabular}

\begin{tabular}[pos]{|l|p{9cm}|}

\hline
Packet ID       &      65
\\ \hline
Name            &      Action
\\ \hline
Size (bytes)    &      2 to 1373
\\ \hline
Format          &      [byte (64)][Action (byte array)]
\\ \hline
Description     &      Used to send text actions to friends. The action is 
                       a non null UTF-8 string with a maximum length of 1372 bytes.
\\ \hline

\end{tabular}

\begin{tabular}[pos]{|l|p{9cm}|}

\hline
Packet ID       &      69
\\ \hline
Name            &      MSI
\\ \hline
Size (bytes)    &      ???
\\ \hline
Format          &      [byte (69)][???]
\\ \hline
Description     &      Used by toxav for initialising audio/video chats.
\\ \hline

\end{tabular}

\begin{tabular}[pos]{|l|p{9cm}|}

\hline
Packet ID       &      80
\\ \hline
Name            &      File Send Request
\\ \hline
Size (bytes)    &      ???
\\ \hline
Format          &      [byte (80)][???]
\\ \hline
Description     &      Used to start a file transfer.
\\ \hline

\end{tabular}

\begin{tabular}[pos]{|l|p{9cm}|}

\hline
Packet ID       &      81
\\ \hline
Name            &      File Send Control
\\ \hline
Size (bytes)    &      ???
\\ \hline
Format          &      [byte (81)][???]
\\ \hline
Description     &      Used to accept, pause, resume and cancel file transfers.
\\ \hline

\end{tabular}

\begin{tabular}[pos]{|l|p{9cm}|}

\hline
Packet ID       &      82
\\ \hline
Name            &      File Data
\\ \hline
Size (bytes)    &      ???
\\ \hline
Format          &      [byte (82)][???]
\\ \hline
Description     &      Used to send actual file data in file transfers.
\\ \hline

\end{tabular}



\end{document}

